Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2017-12-20 PAPER.tex
made jonnie syntax edits throughout paper

\newcommand{\cc}[1]{{\color{purple} \textbf{[CC: #1]}}}
\newcommand{\acl}[1]{{\color{red} \textbf{[ACL:  #1]}}}
\newcommand{\dcj}[1]{{\color{orange} \textbf{[DCJ: #1]}}}
\newcommand{\mjk}[1]{ {\color{olive} \textbf{[MJK: #1]}} }

@@

\author{
Carina Cheng\altaffilmark{1},
Aaron R. Parsons\altaffilmark{1,2},
Matthew Kolopanis \altaffilmark{3},
Daniel C. Jacobs\altaffilmark{3}, 
Saul A. Kohn\altaffilmark{4},
Adrian Liu\altaffilmark{1,5}, 
James E. Aguirre\altaffilmark{4},
Zaki S. Ali\altaffilmark{1}, 
Gianni Bernardi\altaffilmark{6,7,8}, 
Richard F. Bradley\altaffilmark{9,10,11},
Chris L. Carilli\altaffilmark{12,13},
David R. DeBoer\altaffilmark{2}, 
Matthew R. Dexter\altaffilmark{2},
Joshua S. Dillon\altaffilmark{1},
Pat Klima\altaffilmark{7},
David H. E. MacMahon\altaffilmark{2},
David F. Moore\altaffilmark{4},
Chuneeta D. Nunhokee\altaffilmark{10},
Jonathan C. Pober\altaffilmark{14}, 
William P. Walbrugh\altaffilmark{9},
Andre Walker\altaffilmark{9}
}

@@

Bootstrapping uses sampling with replacement to estimate a posterior distribution. For example, measurements (like power 
spectra) can be made from different samples of data. Averaging each sample set gives us a realization of the (power spectrum) 
answer. Realizations are correlated with each other to a degree set by the fraction of sampled points that are held in common 
between them. Hence, through the process of resampling and averaging along different axes, we can estimate error bars for 
our results which represent the underlying distribution of values that are allowed by our measurements.

@@

Because foreground spectra are smooth in frequency, they preferentially show up at low delay, or low $k$ modes. For a particular 
baseline length, there is a maximum delay imposed on foregrounds, which corresponds to the light-crossing time between two 
antennas in a baseline. For longer baselines, this value increases, producing what is known as ``the 
wedge"
\citep{datta_et_al2010, parsons_et_al2012b, vedantham_et_al2012, pober_et_al2013, thyagarajan_et_al2013, liu_et_al2014a, liu_et_al2014b, patil_et_al2017}. 
The wedge describes a region in $k$-space contaminated by foregrounds, bounded by baseline length (which is proportional to $k_{\perp}$) and delay (which is 
proportional to $k_{\parallel}$). Properties of the wedge can be used to isolate and 
%remove 
avoid foregrounds, as done by P64, 
\citet{parsons_et_al2014}, and \citet{jacobs_et_al2015}.

However, this isolation is not perfect.  In deep measurements, power spectrum measurements at $k_{\parallel}$ values beyond 
the length of a baseline are often still contaminated at a low level. This leakage, particularly at low $k$'s, can be attributed to 
convolution kernels associated with Fourier-transforming visibilities into delay-space. In other words, smooth-spectrum 
foregrounds appear as $\delta$-functions in delay-space, convolved by the Fourier transform of the source spectrum and the 
antenna response, both of which could smear out the foregrounds and cause leakage outside the wedge \citep[e.g.][]{ewall-wice_et_al2017}.

There are analysis techniques to mitigate the effects of foreground leakage and prevent information from low $k's$ from 
spreading to high $k$ values. For example, narrow window functions can be used to minimize the covariance of a particular 
$k$ value with other ones (\citealt{liu_et_al2014b}). In other words, one can construct an estimator using OQE that forces a 
window function to have a minimum response to low $k$ values. The window function used in P64 is constructed in such a way, 
specifically to prevent foregrounds that live at low $k's$ from contaminating higher $k$-modes (see Section \ref{sec:Bias}). 
Minimizing foreground leakage in this way however, comes with the trade-off of compromising power spectrum sensitivity, since 
narrow window functions increases errors for each $k$-mode (\citealt{liu_et_al2014b}).

@@

As a brief review, PAPER is a dedicated $21$ cm experiment located in the Karoo Desert in South Africa. The PAPER-64 
configuration consists of 64 dual-polarization drift-scan elements that are arranged in a grid layout. For our case study, we 
focus solely on Stokes I estimated data \citep{moore_et_al2013} from PAPER's $30$ m East/West baselines (Figure 
\ref{fig:ant_layout}). For information about the backend system of PAPER-64, its observations, and data reduction pipeline, we 
refer the reader to \citet{parsons_et_al2010} and P64.

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{plots/ant_layout.png}
	\caption{The PAPER-64 antenna layout. We use only the $30$ m East/West baselines for the revised analysis in this 
paper (such as the baseline involving antennas 49 and 41).}
	\label{fig:ant_layout}
\end{figure}

@@

\begin{equation}
\Delta^{\textbf{2}}\textbf{(k)} = \frac{k^{3}}{2\pi^{2}}\hat{\textbf{p}}\textbf{(k)},
\end{equation}

@@

For our analysis, we use $8$ hours of LST (RA $0.5$-$8.6$ hours) and $51$ total baselines (P64 uses a slightly different RA 
range of $0$-$8.6$ hours). All power spectrum results are produced for a center frequency of $151$ MHz using a width of $10$ 
MHz ($20$ channels), identical to the analysis in P64. We note that, besides using only one baseline type instead of the three in 
P64, the PAPER-64 dataset that we use in this case study differs from that in P64 mainly by the applied fringe-rate filter. In P64, the 
applied filter was degraded by widening it in fringe-rate space. This was chosen in order to increase the number of independent 
modes and reduce signal loss, though as we will explain in the next section, this signal loss was still under-estimated. With the development of a new, 
robust method for assessing signal loss, we choose to use the optimal filter in order to maximize sensitivity. This filter is 
computed for a fiducial $30$ m baseline at $150$ MHz, the center frequency in our band. The filter in both the fringe-rate 
domain and time domain is shown in Figure \ref{fig:frp}.

@@

In Section \ref{sec:SiglossOverview}, we showed how signal loss arises when weighting data using information from the data 
itself. Here we describe a methodology that simulates the injection and recovery of a cosmological signal in order to quantify 
the amount of signal loss accompanying a weighting scheme. In particular, we highlight major differences from the signal loss 
computation used in P64, which previously underestimated losses.

@@

\noindent where we can neglect the first term because $\textbf{x}$ is small.  Without loss of generality, we will assume
an eigenbasis of $\textbf{e}$, so that $\langle 
\textbf{ee}^{\dagger} \rangle_{t}$ is diagonal. 
The middle 
two terms, however, can have power in their off-diagonal terms due to the fact that, when averaging over a finite
ensemble, $\langle\textbf{xe}^\dagger\rangle_t$ is not zero.  As shown in Appendix C of \citet{parsons_et_al2014}, to leading order the
inversion of a matrix like $\hat{\textbf{C}}_{r}$ that is diagonal-dominant (from $\langle 
\textbf{ee}^{\dagger} \rangle_{t}$) with smaller
off-diagonal terms results in a new diagonal-dominant matrix with negated off-diagonal terms. These off-diagonal
terms depend on both $\textbf{x}$ and $\textbf{e}$. When $\hat{\textbf{C}}^{-1}_{r}$ is multiplied into $\textbf{x}$,
the result is a vector that is similar to $\textbf{x}$ but
contains a residual correlation to $\textbf{e}$ from the off-diagonal components of $\hat{\textbf{C}}^{-1}_{r}$. The
correlation is negative because the product $\hat{\textbf{C}}_r^{-1}\textbf{x}$ effectively squares the $\textbf{x}$-dependence
of the off-diagonal terms in $\hat{\textbf{C}}^{-1}_{r}$ while retaining the negative sign that arose from the inversion
of a diagonal-dominant matrix.
% XXX [ARP: we may want to write this all out explicitly in an appendix]

@@

We now shift our attention towards computing signal loss for the fringe-rate filtered PAPER-64 dataset. While our methodology 
outlined below is independent of weighting scheme, here we demonstrate the computation using inverse covariance weighting 
($\textbf{R} = \textbf{C}^{-1}$), the weighting scheme used in P64 which leads to substantial loss. With this weighting, our 
expressions for $P_{\rm in}$ and $P_{\rm out}$ become:

@@

where $\nu$ are frequencies in GHz. We use a receiver temperature of $144$ K, yielding $T_{sys} = 431$ K at $150$ MHz. 
This is lower than the $T_{sys}$ of $500$ K used in P64 because of several small miscalculation errors that were 
identified\footnote{such as missing a square root in going from a variance to a standard deviation}.
\item $\sqrt{2}$: This factor in the denominator of the sensitivity equation comes from taking the real part of the power spectrum 
estimates after squaring visibility measurements. In P64, a factor of $2$ was mistakenly used.
\item $N_{lsts}$: The number of LST hours that go into a power spectrum estimation. The sensitivity scales as the square root 
because we integrate incoherently over time. For PAPER-64, $N_{lsts} = 8$ hours.
\item $N_{seps}$: The number of baseline separation types averaged incoherently in a final power spectrum estimate. For the 
analysis in this paper, we only use one type of baseline, hence $N_{seps}=1$. The updated limits in Kolopanis et al., \textit{in prep}
use three separation types.
\item $t_{int}$: The integration time of the data. It is crucial to adapt this number if filtering is applied along the time axis (i.e. a 
fringe-rate filter). We compute the effective integration time of our fringe-rate filtered data by scaling the original integration time 
using the following:
\begin{equation}
t_{frf} = t_{int} \frac{\int1 \, df}{\int w^{2}(f) \,df},
\end{equation}
where $t_{int}=43$ seconds, $t_{frf}$ is the fringe-rate filtered integration time, $w$ is the fringe-rate profile, and the integral is 
taken over all fringe-rates. For PAPER-64, this number is $t_{int} = 3857$ s. 
\item $N_{days}$: The total number of days of data analyzed. In P64, this number was set to $135$. However, because we 
divide our data in half (to form `even' and `odd' datasets), this number should be reduced by a factor of $2$. Additionally, 
because our LST coverage is not $100\%$ complete (it doesn't overlap for every single day), we compute a more realistic value of 
$N_{days}$ as:
\begin{equation}
 \frac{1}{N_{days}} = \sqrt{\Big\langle\frac{1}{N_{i}^{2}} \Big\rangle_{i}},
 \end{equation}
\noindent where $i$ is over LST (\citealt{jacobs_et_al2015}). For PAPER-64, our revised estimate of $N_{days}$ is $\sim34$ 
days.
\item $N_{bls}$: The number of baselines contributing to the sensitivity of a power spectrum estimate. In P64, this number was 
the total number of $30$ m East/West baselines used in the analysis. However, using the total number of baselines neglects 
the fact that baselines are averaged into groups (see Section \ref{sec:Boot}) before cross-multiplying data. Our revised estimate for the parameter is:
\begin{equation}
N_{bls} = \frac{N_{bls}}{N_{gps}}\sqrt{N_{gps}^{2}-N_{gps}},
\end{equation}
\noindent where, for the P64 analysis, $N_{gps} = 5$. Each baseline group averages down linearly as the number of baselines 
entering the group ($N_{bls}/N_{gps}$) and then as the square root of the number of cross-multiplied pairs ($\sqrt{N_{gps}^{2} - 
N_{gps}}$). For the revised PAPER-64 analysis with only one baseline separation type, this becomes $N_{bls} \sim 46$ instead 
of $51$. 
\item $N_{pols}$: The number of polarizations averaged together. For the case of Stokes I, $N_{pols}=2$.
\end{itemize}

@@

\begin{itemize}
\item Substantial signal loss can result when weighting data using empirically estimated covariances (Section 
\ref{sec:SiglossOverview}). Loss of the $21$ cm signal is especially significant if fewer modes
exist in the data. Hence, there exists a trade-off between sensitivity driven 
time-averaging techniques such as fringe-rate filtering and signal loss. 
\item Signal injection and recovery simulations can be used to quantify signal loss (Section \ref{sec:Sigloss}). However, a 
signal-only simulation (i.e. comparing an unweighted vs. weighted power spectrum of EoR only) can underestimate loss by 
failing to account for correlations between the data and signal (Section \ref{sec:Illustration}).
\item Errors that are estimated via bootstrapping can be underestimated if samples in the dataset are significantly correlated 
(Section \ref{sec:ErrorOverview}). If the number of independent samples in a dataset is well-determined, bootstrapping is a 
simple and accurate way of estimating errors.
\item Meaningful null tests are vital to validate an EoR detection (Section \ref{sec:JackknifeOverview}). Similarly, performing 
jackknife tests along multiple axes of a dataset is necessary for confidence in an EoR detection and can also be used to tease 
out systematics.
\end{itemize}

@@

In this Appendix, we examine a toy model for signal loss. Our goal is to build intuition by deriving an analytic formula for power spectrum signal loss. We will also show that in general, signal loss appears as a multiplicative bias on one's power spectrum estimate.

@@

In the main text we made an estimate of the signal loss effecting only the injected EoR by supposing that the lossy injected signal was the difference between the injected and not injected cases: $P_{out} = \hat{\textbf{p}}_{r} - \hat{\textbf{p}}_{x}$. Suppose that we rephrase the question we are asking to: ``At what point would an injected signal be detectable at significant levels?". Instead of detangling the part of the output power spectrum dominated by the injected signal we can simply use $\hat{\textbf{p}}_{r}$, the weighted power spectrum of the data plus injected signal.

@@

\begin{figure}[tp]
\centering
\includegraphics[width=.5\textwidth]{plots/method2_prob_vs_pinj.png}
\caption{The probability that an injected signal of level P$_{inj}$
is observable at a 2$\sigma$ level above the value of $\hat{\textbf{p}}_{x}$ (data).
The different colors represent different $k$ values. We choose a probably of $90\%$ for our signal loss estimated power spectrum.
\label{fig:Prob_vs_Pin}}
\end{figure}

>>>>

\newcommand{\cc}[1]{{\color{purple} \textbf{[CC: #1]}}}
\newcommand{\acl}[1]{{\color{red} \textbf{[ACL:  #1]}}}
\newcommand{\dcj}[1]{{\color{orange} \textbf{[DCJ: #1]}}}
\newcommand{\mjk}[1]{ {\color{olive} \textbf{[MJK: #1]}} }
\newcommand{\jea}[1]{{\color{blue} \textbf{[JEA: #1]}} }

@@

\author{
Carina Cheng\altaffilmark{1},
Aaron R. Parsons\altaffilmark{1,2},
Matthew Kolopanis \altaffilmark{3},
Daniel C. Jacobs\altaffilmark{3}, 
Saul A. Kohn\altaffilmark{4},
Adrian Liu\altaffilmark{1,5}, 
James E.~Aguirre\altaffilmark{4},
Zaki S. Ali\altaffilmark{1}, 
Gianni Bernardi\altaffilmark{6,7,8}, 
Richard F. Bradley\altaffilmark{9,10,11},
Chris L. Carilli\altaffilmark{12,13},
David R. DeBoer\altaffilmark{2}, 
Matthew R. Dexter\altaffilmark{2},
Joshua S. Dillon\altaffilmark{1},
Pat Klima\altaffilmark{7},
David H. E. MacMahon\altaffilmark{2},
David F. Moore\altaffilmark{4},
Chuneeta D. Nunhokee\altaffilmark{10},
Jonathan C. Pober\altaffilmark{14}, 
William P. Walbrugh\altaffilmark{9},
Andre Walker\altaffilmark{9}
}

@@

Bootstrapping uses sampling with replacement to estimate a posterior distribution. For example, measurements (like power 
spectra) can be made from different samples of data. Each of these measurements is a different realization drawn from some underlying distribution, and realizations are correlated with each other to a degree set by the fraction of sampled points that are held in common 
between them. Hence, through the process of resampling and averaging along different axes, we can estimate error bars for 
our results which represent the underlying distribution of values that are allowed by our measurements.

@@

For a particular 
baseline length, there is a maximum delay imposed on sources attached to the sky, which corresponds to the light-crossing time between two 
antennas in a baseline. For longer baselines, this value increases, producing what is known as ``the 
wedge"
\citep{datta_et_al2010, parsons_et_al2012b, vedantham_et_al2012, pober_et_al2013, thyagarajan_et_al2013, liu_et_al2014a, liu_et_al2014b, patil_et_al2017}. 
The wedge describes a region in $k$-space contaminated by smooth spectrum foregrounds, bounded by baseline length (which is proportional to $k_{\perp}$) and delay (which is 
proportional to $k_{\parallel}$). Properties of the wedge can be used to isolate and 
%remove 
avoid foregrounds, as done by P64, 
\citet{parsons_et_al2014}, and \citet{jacobs_et_al2015}.

Although smooth spectrum foregrounds preferentially show up at low delay, or low $k$ modes, their isolation within the wedge is not perfect.  In deep measurements, power spectrum measurements at $k_{\parallel}$ values beyond 
the length of a baseline are often still contaminated at a low level. This leakage, particularly at low $k$'s, can be attributed to 
convolution kernels associated with Fourier-transforming visibilities into delay-space. In other words, smooth-spectrum 
foregrounds appear as $\delta$-functions in delay-space, convolved by the Fourier transform of the source spectrum and the 
antenna response, both of which could smear out the foregrounds and cause leakage outside the wedge \citep[e.g.][]{ewall-wice_et_al2017}.

There are analysis techniques to mitigate the effects of foreground leakage and prevent information from low $k's$ from 
spreading to high $k$ values. For example, narrow window functions in delay-space can be used to minimize the covariance of a particular 
$k$ value with other ones (\citealt{liu_et_al2014b}). In other words, one can construct an estimator using OQE that forces a 
window function to have a minimum response to low $k$ values. The window function used in P64 is constructed in such a way, 
specifically to prevent foregrounds that live at low $k's$ from contaminating higher $k$-modes (see Section \ref{sec:Bias}). 
Minimizing foreground leakage in this way however, comes with the trade-off of compromising power spectrum sensitivity, since 
narrow window functions increases errors for each $k$-mode (\citealt{liu_et_al2014b}).

@@

As a brief review, PAPER is a dedicated $21$ cm experiment located in the Karoo Desert in South Africa. The PAPER-64 
configuration consists of 64 dual-polarization drift-scan elements that are arranged in a grid layout. For our case study, we 
focus solely on Stokes I estimated data \citep{moore_et_al2013} from PAPER's $30$ m East/West baselines (Figure 
\ref{fig:ant_layout}). All data is compressed, calibrated (using self-calibration and redundant calibration), delay-filtered (to remove foregrounds inside the wedge), LST-binned, and fringe-rate filtered. For detailed information about the backend system of PAPER-64, its observations, and data reduction pipeline, we 
refer the reader to \citet{parsons_et_al2010} and P64. We note that all data processing steps are identical to those in P64 until after the LST-binning step in Figure 3 of P64.

\begin{figure}
	\centering
	\includegraphics[trim={0cm 4cm 0cm 5cm},width=\columnwidth]{plots/ant_layout_aspect.png}
	\caption{The PAPER-64 antenna layout. We use only the $30$ m East/West baselines for the revised analysis in this 
paper (i.e. the shortest horizontal spacings).}
	\label{fig:ant_layout}
\end{figure}

@@

\begin{equation}
\Delta^{\textbf{2}}(k) = \frac{k^{3}}{2\pi^{2}}\hat{\textbf{p}}\textbf{(k)},
\end{equation}

@@

For our analysis, we use $8.1$ hours of LST (RA $0.5$-$8.6$ hours) and $51$ total baselines (P64 uses a slightly different RA 
range of $0$-$8.6$ hours). All power spectrum results are produced for a center frequency of $151$ MHz using a width of $10$ 
MHz ($20$ channels), identical to the analysis in P64. We note that, besides using only one baseline type instead of the three in 
P64, the PAPER-64 dataset that we use in this case study differs from that in P64 mainly by the applied fringe-rate filter. In P64, the 
applied filter was degraded by widening it in fringe-rate space. This was chosen in order to increase the number of independent 
modes and reduce signal loss, though as we will explain in the next section, this signal loss was still under-estimated. With the development of a new, 
robust method for assessing signal loss, we choose to use the optimal filter in order to maximize sensitivity. This filter is 
computed for a fiducial $30$ m baseline at $150$ MHz, the center frequency in our band. The filter in both the fringe-rate 
domain and time domain is shown in Figure \ref{fig:frp}.

@@

In Section \ref{sec:SiglossOverview}, we showed how signal loss arises when weighting data using information from the data 
itself. Here we describe a methodology that simulates the injection and recovery of a cosmological signal in order to quantify 
the amount of signal loss accompanying a weighting scheme used in OQE power spectrum estimation. In particular, we highlight major differences from the signal loss 
computation used in P64, which previously underestimated losses.

@@

\noindent where we can neglect the first term because $\textbf{x}$ is small.  Without loss of generality, we will assume
an eigenbasis of $\textbf{e}$, so that $\langle 
\textbf{ee}^{\dagger} \rangle_{t}$ is diagonal. 
The middle 
two terms, however, can have power in their off-diagonal terms due to the fact that, when averaging over a finite
ensemble, $\langle\textbf{xe}^\dagger\rangle_t$ is not zero.  As shown in Appendix C of \citet{parsons_et_al2014}, to leading order the
inversion of a matrix like $\hat{\textbf{C}}_{r}$ that is diagonal-dominant (from $\langle 
\textbf{ee}^{\dagger} \rangle_{t}$) with smaller
off-diagonal terms results in a new diagonal-dominant matrix with negative off-diagonal terms. These off-diagonal
terms depend on both $\textbf{x}$ and $\textbf{e}$. When $\hat{\textbf{C}}^{-1}_{r}$ is multiplied into $\textbf{x}$,
the result is a vector that is similar to $\textbf{x}$ but
contains a residual correlation to $\textbf{e}$ from the off-diagonal components of $\hat{\textbf{C}}^{-1}_{r}$. The
correlation is negative because the product $\hat{\textbf{C}}_r^{-1}\textbf{x}$ effectively squares the $\textbf{x}$-dependence
of the off-diagonal terms in $\hat{\textbf{C}}^{-1}_{r}$ while retaining the negative sign that arose from the inversion
of a diagonal-dominant matrix.
% XXX [ARP: we may want to write this all out explicitly in an appendix]

@@

We now shift our attention towards computing signal loss for the fringe-rate filtered PAPER-64 dataset. While our methodology 
outlined below is independent of weighting scheme, here we demonstrate the computation using inverse covariance weighting 
($\textbf{R} = \hat{\textbf{C}}^{-1}$), the weighting scheme used in P64 which, with an empirically estimated covariance, leads to substantial loss. With this weighting, our 
expressions for $P_{\rm in}$ and $P_{\rm out}$ become:

@@

where $\nu$ are frequencies in GHz. We use a receiver temperature of $144$ K, yielding $T_{sys} = 431$ K at $150$ MHz. 
This is lower than the $T_{sys}$ of $500$ K used in P64 because of several small miscalculation errors that were 
identified\footnote{such as missing a square root in going from a variance to a standard deviation}.
\item $\sqrt{2}$: This factor in the denominator of the sensitivity equation comes from taking the real part of the power spectrum 
estimates after squaring visibility measurements. In P64, a factor of $2$ was mistakenly used.
\item $N_{lsts}$: The number of LST hours that go into a power spectrum estimation. The sensitivity scales as the square root 
because we integrate incoherently over time. For PAPER-64, $N_{lsts} = 8$ hours.
\item $N_{seps}$: The number of baseline separation types averaged incoherently in a final power spectrum estimate. For the 
analysis in this paper, we only use one type of baseline, hence $N_{seps}=1$. The updated limits in Kolopanis et al., \textit{in prep}
use three separation types.
\item $t_{int}$: The integration time of the data. It is crucial to adapt this number if filtering is applied along the time axis (i.e. a 
fringe-rate filter). We compute the effective integration time of our fringe-rate filtered data by scaling the original integration time $t_{i}$
using the following:
\begin{equation}
t_{int} = t_{i} \frac{\int1 \, df}{\int w^{2}(f) \,df},
\end{equation}
where $t_{i}=43$ seconds, $t_{int}$ is the fringe-rate filtered integration time, $w$ is the fringe-rate profile, and the integral is 
taken over all fringe-rates. For PAPER-64, this number is $t_{int} = 3857$ s. 
\item $N_{days}$: The total number of days of data analyzed. In P64, this number was set to $135$. However, because we 
divide our data in half (to form `even' and `odd' datasets), this number should be reduced by a factor of $2$. Additionally, 
because our LST coverage is not $100\%$ complete (it doesn't overlap for every single day), we compute a more realistic value of 
$N_{days}$ as:
\begin{equation}
 \frac{1}{N_{days}} = \sqrt{\Big\langle\frac{1}{N_{i}^{2}} \Big\rangle_{i}},
 \end{equation}
\noindent where $i$ is over LST (\citealt{jacobs_et_al2015}). For PAPER-64, our revised estimate of $N_{days}$ is $\sim34$ 
days.
\item $N_{bls}$: The number of baselines contributing to the sensitivity of a power spectrum estimate. In P64, this number was 
the total number of $30$ m East/West baselines used in the analysis. However, using the total number of baselines neglects 
the fact that baselines are averaged into groups (see Section \ref{sec:Boot}) before cross-multiplying data. Our revised estimate for the parameter is:
\begin{equation}
N_{bls} = \frac{N_{bls}}{N_{gps}}\sqrt{N_{gps}^{2}-N_{gps}},
\end{equation}
\noindent where, for the P64 analysis, $N_{gps} = 5$. Each baseline group averages down linearly as the number of baselines 
entering the group ($N_{bls}/N_{gps}$) and then as the square root of the number of cross-multiplied pairs ($\sqrt{N_{gps}^{2} - 
N_{gps}}$). For the revised PAPER-64 analysis with only one baseline separation type, this becomes $N_{bls} \sim 46$ instead 
of $51$. 
\item $N_{pols}$: The number of polarizations averaged together. For the case of Stokes I, $N_{pols}=2$.
\end{itemize}

@@

\begin{itemize}
\item Substantial signal loss can result when weighting data using empirically estimated covariances (Section 
\ref{sec:SiglossOverview}). Loss of the $21$ cm signal is especially significant the fewer number of independent modes that
exist in the data. Hence, there exists a trade-off between sensitivity driven 
time-averaging techniques such as fringe-rate filtering and signal loss when using empirically estimated covariances. 
\item Signal injection and recovery simulations can be used to quantify signal loss (Section \ref{sec:Sigloss}). However, a 
signal-only simulation (i.e. comparing an unweighted vs. weighted power spectrum of EoR only) can underestimate loss by 
failing to account for correlations between the data and signal (Section \ref{sec:Illustration}).
\item Errors that are estimated via bootstrapping can be underestimated if samples in the dataset are significantly correlated 
(Section \ref{sec:ErrorOverview}). If the number of independent samples in a dataset is well-determined, bootstrapping is a 
simple and accurate way of estimating errors.
\item Meaningful null tests are vital to validate an EoR detection (Section \ref{sec:JackknifeOverview}). Similarly, performing 
jackknife tests along multiple axes of a dataset is necessary for confidence in an EoR detection and can also be used to tease 
out systematics.
\end{itemize}

@@

In this Appendix, we examine a toy model for signal loss. Our goal is to derive an analytic formula for power spectrum signal loss. We will also show that in general, signal loss appears as a multiplicative bias on one's power spectrum estimate.

@@

In the main text we made an estimate of the signal loss effecting only the injected EoR by supposing that the lossy injected signal was the difference between the injected and not injected cases: $P_{out} = \hat{\textbf{p}}_{r} - \hat{\textbf{p}}_{x}$. Suppose that we rephrase the question we are asking to: ``At what point would an injected signal be detectable at significant levels?". Instead of disentangling the part of the output power spectrum dominated by the injected signal we can simply use $\hat{\textbf{p}}_{r}$, the weighted power spectrum of the data plus injected signal.

@@

\begin{figure}[tp]
\centering
\includegraphics[width=.5\textwidth]{plots/method2_prob_vs_pinj.png}
\caption{The probability that an injected signal of level P$_{inj}$
is observable at a 2$\sigma$ level above the value of $\hat{\textbf{p}}_{x}$ (data).
The different colors represent different $k$ values. We choose a probability of $95\%$ for our signal loss estimated power spectrum.
\label{fig:Prob_vs_Pin}}
\end{figure}
